---
title: "KB PCV"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages('survminer')
library(lubridate)
library(zoo)
library(reshape2)
library(survival)
library(survminer)
library(htmlTable)
library(polySegratio)
```

## Rollout of PCV in India
We are using data from Kushi baby to evaluate the rollout of PCVs and timeliness of vaccination, and considering the potential for PCTS data to evaluate effect on pneumonia hospitalizations


```{r import, echo=FALSE}
ds1<-read.csv('../kb_data/KB_Child_Report.csv')

```

#Vaccine rollout among KB population
What should be denominator here? Ideally would be # of kids who are eligible to receive vaccine in that district. Can look at kids in database who have birthdate after 
Shows that May 2018 is when Number of doses delivered started ot pick up (spike in September). A few kids have vaccine receipt dates that are earlier. Are these real, or are they data entry errors?
-It also looks like there must be a catch up campaign in effect, since there were a number of kids who received doses at older ages than would be expected if everyone was being vaccinated according to an EPI schedule.
```{r rollout, echo=FALSE}
pcv1<-rep(0, times=nrow(ds1))
pcv2<-rep(0, times=nrow(ds1))
pcv3<-rep(0, times=nrow(ds1))

pcv1[ds1$PCV1.Given.Denied=='Given']<-1 
pcv2[ds1$PCV2.Given.Denied=='Given']<-1 
pcv3[ds1$PCV.Booster.Given.Denied=='Given']<-1 

print("Number of kids with 1,2,3 doses")
sum(pcv1, na.rm=TRUE)
sum(pcv2, na.rm=TRUE)
sum(pcv3, na.rm=TRUE)

```

## When were first doses distributed?
```{r}
hist(as.Date(ds1$PCV1.Date[ds1$PCV1.Given.Denied =='Given']), 'month', main='Date of First dose receipt', xlab='Date')

dob_month<-as.Date(as.yearmon(as.Date(ds1$DOB)))
par(mfrow=c(1,2))
hist(dob_month, 'month', main='DOB of all kids in Kushi baby cohort')
#Look at birthdays of kids who got 1 dose--have they been eligible for 4 month and 9 month doses yet?
dob.vaccinated.dose1<-as.Date(ds1$DOB[which(pcv1==1)])

hist(dob.vaccinated.dose1, 'months', main='DOB of kids who received 1 dose of vaccine')
dob.vaccinated.dose2<-as.Date(ds1$DOB[which(pcv2==1)])
par(mfrow=c(1,1))
```

#Coverage by birth cohort
Since it looks like vaccination took off in earnest in May 2018, we will consider all kids born on or after March 1, 2018.
```{r epi.only, echo=FALSE}
routine.eligible.flag<-rep(0, times=nrow(ds1))
routine.eligible.flag[as.Date(ds1$DOB)>=as.Date('2018-03-01') ]<-1
#table(routine.eligible.flag, pcv1)

#tab1<-table(dob_month[routine.eligible.flag==1], pcv1[routine.eligible.flag==1])
tab1<-table(dob_month, pcv1)
tab1.prop<-prop.table(tab1, margin = 1)

tab2<-table(dob_month, pcv2)
tab2.prop<-prop.table(tab2, margin = 1)

tab3<-table(dob_month, pcv3)
tab3.prop<-prop.table(tab3, margin = 1)

monthdates<-unique(floor_date(as.Date(ds1$DOB,"%Y-%m-%d" ), unit='month' ) ) 
monthdates<-monthdates[!is.na(monthdates)]

monthdates<-monthdates[monthdates %in% as.Date(dimnames(tab2.prop)[[1]]) ]
monthdates<-sort(monthdates)

print("Number of eligible kids registered in KB system who did not or did receive first dose of PCV, based on Month of birth")
print(tab1)

print("Proportion of eligible kids registered in KB system who did not or did receive first dose of PCV, based on month of birth")
print(round(tab1.prop,2))

plot(monthdates,100*round(tab1.prop,2)[,2], ylab='Percent vaccinated', xlab="Birth Month", main='Uptake by dose and birth month as of Oct 2019', bty='l', ylim=c(0, 100), type='l', xaxt='n', xlim=as.Date(c('2016-07-01','2019-10-01')))
points(monthdates,100*round(tab2.prop,2)[,2], type='l', col='blue')
points(monthdates,100*round(tab3.prop,2)[,2], type='l', col='red')
axis(side=1, at=as.Date(c('2016-07-01','2017-07-01','2018-03-01','2018-09-01', '2019-03-01', '2019-09-01')), labels=c('Jul 2016','Jul 2017','Mar 2018','Sep 2018', 'Mar 2019', 'Sep 2019'))
```


What is rate of refusal?
```{r}

table(ds1$PCV1.Given.Denied[routine.eligible.flag==1])
table(ds1$PCV2.Given.Denied[routine.eligible.flag==1])
table(ds1$PCV.Booster.Given.Denied[routine.eligible.flag==1])

```

# Look at uptake of 1,2, 3 doses by age
First look at  uptake of 1,2, or booster dose by age in all eligible kids born after March 1,2018. his is done as a survival model because of the structure of the data. Many kids in the database don't become age-eligible to receive a dose, so we censor them.
```{r}

long1<-ds1[,c('Child.Registration.ID', 'DOB', 'PCV1.Date', 'PCV1.Given.Denied', 'PCV2.Date', 'PCV2.Given.Denied','PCV.Booster.Date', 'PCV.Booster.Given.Denied' )]

long1$PCV1.Date<-as.Date(long1$PCV1.Date, format='%Y-%m-%d')
long1$PCV2.Date<-as.Date(long1$PCV2.Date, format='%Y-%m-%d')
long1$PCV.Booster.Date<-as.Date(long1$PCV.Booster.Date, format='%Y-%m-%d')
long1$DOB<-as.Date(long1$DOB, format='%Y-%m-%d')

long1$PCV1.Date[long1$PCV1.Given.Denied!='Given']<-NA
long1$PCV2.Date[long1$PCV2.Given.Denied!='Given']<-NA
long1$PCV.Booster.Date[long1$PCV.Booster.Given.Denied!='Given']<-NA
long1$censor.date<-max(long1$PCV1.Date, na.rm=TRUE)

long1$pcv1.aged<- as.numeric(long1$PCV1.Date - long1$DOB)
long1$pcv2.aged<- as.numeric(long1$PCV2.Date - long1$DOB)
long1$pcv.boost.aged<- as.numeric(long1$PCV.Booster.Date - long1$DOB)
long1$censor.age<-as.numeric(long1$censor.date - long1$DOB)
long1$init.age=0
long1<-long1[long1$DOB>=as.Date('2018-03-01'),] #only eligible kids

long2<-long1[,c('Child.Registration.ID', 'pcv1.aged',  'pcv2.aged', 'pcv.boost.aged','censor.age','init.age')]
long2.m<-melt(long2, id.var=c('Child.Registration.ID'))
long2.c<-dcast(long2.m,Child.Registration.ID+variable~. )
names(long2.c) <- c('id','variable','aged')
long2.c<-long2.c[!is.na(long2.c$aged),]
 
format.ds<-function(ds, var1){
  doseX.tj<-ds[ds$variable %in% c('censor.age','init.age',var1) ,]
  doseX.tj$id<-factor(doseX.tj$id)
  doseX.tj.spl<-split(doseX.tj, doseX.tj$id)
  doseX.tj.spl<-lapply(doseX.tj.spl, function(x){
    x$doseX<-0
    x$doseX[x$variable==var1]<-1
    x$doseXN<-sum(x$doseX)
    x<- x[x$doseXN==0 | (x$doseXN==1 & x$variable != 'censor.age'),]
    return(x)
  }  )
  doseX.tj<-do.call('rbind.data.frame', doseX.tj.spl)
  doseX.tj$aged[doseX.tj$aged>(365*5)] <- 365*5
  return(doseX.tj)
}
dose1.clean<-format.ds(long2.c, 'pcv1.aged')
dose1.clean$dose=1
dose2.clean<-format.ds(long2.c, 'pcv2.aged')
dose2.clean$dose=2
dose3.clean<-format.ds(long2.c, 'pcv.boost.aged')
dose3.clean$dose<-3

dose.comp<-rbind.data.frame(dose1.clean,dose2.clean,dose3.clean)
dose.comp$agem<-dose.comp$aged/30.3
dose.comp$agem[dose.comp$agem<0]<-0

dose1.mod<- survfit(Surv(dose.comp$agem,dose.comp$doseX)~dose, data=dose.comp)
ggsurvplot(dose1.mod, data = dose.comp, pval =F,  fun = "event", xlab='Age (months)', ylab='Proportion vaccinated')

surv.summary<-surv_summary(dose1.mod,data=dose.comp)
surv.summary.spl<-split(surv.summary, surv.summary$dose)
uptake<-1-sapply(surv.summary.spl, function(x) x$surv[length(x$surv)])
uptake
#summary(dose1.mod)$table
#matrix(summary(dose1.mod)$surv, ncol=3)
#1:201
#202:452
#453:571

plot1<-ggsurvplot(dose1.mod, data = dose.comp, pval =F,  fun = "event", xlab='Age (months)', ylab='Proportion vaccinated')
#plot1+ geom_vline(xintercept = 1:5)

```
Timeliness: define based on within 4 weeks of recomended date https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5720480/
recommended is 1.5, 3.5 and 9 months
```{r}
surv.summary.spl<-split(surv.summary, surv.summary$dose)
rec.month<-c(1.5,3.5,9)
cov.1.month.after.rec<-rep(NA,3)
cov.2.month.after.rec<-rep(NA,3)

for(i in 1:3){
  ds<-surv.summary.spl[[i]]
  #index for 1 months after recommended date
    rec.index1<-which( (abs(ds$time-(rec.month[i]+1)))==min(abs(ds$time-(rec.month[i]+1))))
  rec.index2<-which( (abs(ds$time-(rec.month[i]+2)))==min(abs(ds$time-(rec.month[i]+2))))
  cov.1.month.after.rec[i]<- 1-ds$surv[rec.index1]
  cov.2.month.after.rec[i]<- 1-ds$surv[rec.index2]

}
timeliness<-cbind.data.frame('doses'=c(1,2,3),'uptake'=round(uptake,2),'cov.1m'=round(cov.1.month.after.rec,2),
'cov.2m'=round(cov.2.month.after.rec,2), 'pct.timely1m'=round(cov.1.month.after.rec/uptake,2), 'pct.timely2m'=round(cov.2.month.after.rec/uptake,2))
htmlTable(timeliness)

```




#Spatial variability in uptake
```{r, fig.width=7, fig.height=2.5}
unique(ds1$Block)
#unique(ds1$PHC)

totalN.block<-as.matrix(table(ds1$Block))
N.block.dob.march2018<-as.matrix(table(ds1$Block[as.Date(ds1$DOB)>=as.Date('2018-01-01')] ))
blocknames<-row.names(totalN.block)
tab1<-cbind.data.frame(totalN.block,N.block.dob.march2018)
tab1<-rbind(tab1, apply(tab1,2,sum))
tab1<-cbind.data.frame( c(blocknames,'Total'), tab1)
names(tab1)<-c('Admin. Region', 'All Children', 'Children born after Mar 2018')
row.names(tab1)<-NULL
htmlTable(tab1, rnames=F,align='c',caption="Table 1. Number of children in the Khushi Baby database, October 2019")

long1<-ds1[,c('Child.Registration.ID', 'DOB', 'PCV1.Date', 'PCV1.Given.Denied', 'PCV2.Date', 'PCV2.Given.Denied','PCV.Booster.Date', 'PCV.Booster.Given.Denied','Block' )]

long1$PCV1.Date<-as.Date(long1$PCV1.Date, format='%Y-%m-%d')
long1$PCV2.Date<-as.Date(long1$PCV2.Date, format='%Y-%m-%d')
long1$PCV.Booster.Date<-as.Date(long1$PCV.Booster.Date, format='%Y-%m-%d')
long1$DOB<-as.Date(long1$DOB, format='%Y-%m-%d')

long1$PCV1.Date[long1$PCV1.Given.Denied!='Given']<-NA
long1$PCV2.Date[long1$PCV2.Given.Denied!='Given']<-NA
long1$PCV.Booster.Date[long1$PCV.Booster.Given.Denied!='Given']<-NA
long1$censor.date<-max(long1$PCV1.Date, na.rm=TRUE)

long1$pcv1.aged<- as.numeric(long1$PCV1.Date - long1$DOB)
long1$pcv2.aged<- as.numeric(long1$PCV2.Date - long1$DOB)
long1$pcv.boost.aged<- as.numeric(long1$PCV.Booster.Date - long1$DOB)
long1$censor.age<-as.numeric(long1$censor.date - long1$DOB)
long1$init.age=0
long1<-long1[long1$DOB>=as.Date('2018-03-01'),] #only eligible kids

long2<-long1[,c('Child.Registration.ID', 'pcv1.aged',  'pcv2.aged', 'pcv.boost.aged','censor.age','init.age','Block')]
long2.m<-melt(long2, id.var=c('Child.Registration.ID','Block'))
long2.c<-dcast(long2.m,Child.Registration.ID+variable+Block~. )
names(long2.c) <- c('id','variable','Block','aged')
long2.c<-long2.c[!is.na(long2.c$aged),]
 
format.ds<-function(ds, var1, doseN){
  doseX.tj<-ds[ds$variable %in% c('censor.age','init.age',var1) ,]
  doseX.tj$id<-factor(doseX.tj$id)
  doseX.tj.spl<-split(doseX.tj, doseX.tj$id)
  doseX.tj.spl<-lapply(doseX.tj.spl, function(x){
    x$doseX<-0
    x$doseX[x$variable==var1]<-1
    x$doseXN<-sum(x$doseX)
    x<- x[x$doseXN==0 | (x$doseXN==1 & x$variable != 'censor.age'),]
    return(x)
  }  )
  doseX.tj<-do.call('rbind.data.frame', doseX.tj.spl)
  doseX.tj$aged[doseX.tj$aged>(365*5)] <- 365*5
  doseX.tj$doseN<-doseN
  return(doseX.tj)
}
long2.c.spl<-split(long2.c, long2.c$Block)
dose1.clean<-lapply(long2.c.spl, format.ds, var1='pcv1.aged', doseN=1)
dose2.clean<-lapply(long2.c.spl, format.ds, var1='pcv2.aged', doseN=2)
dose3.clean<-lapply(long2.c.spl, format.ds, var1='pcv.boost.aged', doseN=3)
dose1.clean<-do.call('rbind.data.frame', dose1.clean)
dose2.clean<-do.call('rbind.data.frame', dose2.clean)
dose3.clean<-do.call('rbind.data.frame', dose3.clean)

dose.comp<-rbind.data.frame(dose1.clean,dose2.clean,dose3.clean)
dose.comp$agem<-dose.comp$aged/30.3
dose.comp$agem[dose.comp$agem<0]<-0

dose.comp<-dose.comp[order(dose.comp$Block,dose.comp$doseN),]

dose.comp$Block<-factor(dose.comp$Block)
blocks<-unique(dose.comp$Block)

cols<-c('#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00')
dosemod<- survfit(Surv(dose.comp$agem,dose.comp$doseX)~doseN+Block, data=dose.comp)
#ggsurvplot(dosemod, data = dose.comp, pval =F,  fun = "event",keep.data = TRUE, color='Block', palette=cols)

surv.summary<-surv_summary(dosemod,data=dose.comp)
surv.summary.spl<-split(surv.summary, paste(surv.summary$doseN, surv.summary$Block))
uptake<-1-sapply(surv.summary.spl, function(x) x$surv[length(x$surv)])
uptake
#plot(surv.summary$time, 1-surv.summary$surv, col=cols[surv.summary$Block],bty='l', xlab='Age, months', ylab='Proportion vaccinated')
#abline(h=c(0.2,0.4,0.6,0.8), col='lightgray')

surv.summary.c<-1-acast(surv.summary[,c('time','doseN','Block','surv') ], time~doseN~Block )
time.m<-as.numeric(dimnames(surv.summary.c)[[1]])
target.time<-c(1.5,3.5, 9) #when should the vaccine be given?
par(mfrow=c(1,3), mar=c(4,4,1,1))
for(i in 1:3){
  matplot(time.m,surv.summary.c[,i,], type='p', cex=0.25, pch=16,col=cols,bty='l', xlab='Age, months', ylab='Proportion vaccinated',lty=1, ylim=c(0,1))
  abline(v=target.time[i], col='gray', lty=2)
}

#dosemod2<- survfit(Surv(dose.comp$agem,dose.comp$doseX)~Block, data=dose.comp)
#ggsurv.list <- ggsurvplot_group_by(dosemod2, dose.comp, group.by = c("doseN"))
```

## Create function to run analysis by subgroup
Do not need to modify this function. Call it in the chunk below
```{r}
stratified.function<-function(Select.Var){
    Var.Levels<-unique(ds1[,Select.Var])
    totalN.Strata<-as.matrix(table(ds1[,Select.Var]))
    N.Strata.dob.march2018<-as.matrix(table(ds1[as.Date(ds1$DOB)>=as.Date('2018-01-01'), Select.Var] ))
    Stratanames<-row.names(totalN.Strata)
    tab1<-cbind.data.frame(totalN.Strata,N.Strata.dob.march2018)
    tab1<-rbind(tab1, apply(tab1,2,sum))
    tab1<-cbind.data.frame( c(Stratanames,'Total'), tab1)
    names(tab1)<-c('Admin. Region', 'All Children', 'Children born after Mar 2018')
    row.names(tab1)<-NULL
    strata.tab1<-htmlTable(tab1, rnames=F,align='c',caption="Table 1. Number of children in the Khushi Baby database, October 2019")
    
    long1<-ds1[,c('Child.Registration.ID', 'DOB', 'PCV1.Date', 'PCV1.Given.Denied', 'PCV2.Date', 'PCV2.Given.Denied','PCV.Booster.Date', 'PCV.Booster.Given.Denied',Select.Var )]
    
    long1$PCV1.Date<-as.Date(long1$PCV1.Date, format='%Y-%m-%d')
    long1$PCV2.Date<-as.Date(long1$PCV2.Date, format='%Y-%m-%d')
    long1$PCV.Booster.Date<-as.Date(long1$PCV.Booster.Date, format='%Y-%m-%d')
    long1$DOB<-as.Date(long1$DOB, format='%Y-%m-%d')
    
    long1$PCV1.Date[long1$PCV1.Given.Denied!='Given']<-NA
    long1$PCV2.Date[long1$PCV2.Given.Denied!='Given']<-NA
    long1$PCV.Booster.Date[long1$PCV.Booster.Given.Denied!='Given']<-NA
    long1$censor.date<-max(long1$PCV1.Date, na.rm=TRUE)
    
    long1$pcv1.aged<- as.numeric(long1$PCV1.Date - long1$DOB)
    long1$pcv2.aged<- as.numeric(long1$PCV2.Date - long1$DOB)
    long1$pcv.boost.aged<- as.numeric(long1$PCV.Booster.Date - long1$DOB)
    long1$censor.age<-as.numeric(long1$censor.date - long1$DOB)
    long1$init.age=0
    long1<-long1[long1$DOB>=as.Date('2018-03-01'),] #only eligible kids
    
    long2<-long1[,c('Child.Registration.ID', 'pcv1.aged',  'pcv2.aged', 'pcv.boost.aged','censor.age','init.age',Select.Var)]
    long2.m<-melt(long2, id.var=c('Child.Registration.ID',Select.Var))
    form1<-as.formula(paste0('Child.Registration.ID+variable+', Select.Var, '~.'))
    long2.c<-dcast(long2.m,form1 )
    names(long2.c) <- c('id','variable',Select.Var,'aged')
    long2.c<-long2.c[!is.na(long2.c$aged),]
     
    format.ds<-function(ds, var1, doseN){
      doseX.tj<-ds[ds$variable %in% c('censor.age','init.age',var1) ,]
      doseX.tj$id<-factor(doseX.tj$id)
      doseX.tj.spl<-split(doseX.tj, doseX.tj$id)
      doseX.tj.spl<-lapply(doseX.tj.spl, function(x){
        x$doseX<-0
        x$doseX[x$variable==var1]<-1
        x$doseXN<-sum(x$doseX)
        x<- x[x$doseXN==0 | (x$doseXN==1 & x$variable != 'censor.age'),]
        return(x)
      }  )
      doseX.tj<-do.call('rbind.data.frame', doseX.tj.spl)
      doseX.tj$aged[doseX.tj$aged>(365*2)] <- 365*2
      doseX.tj$doseN<-doseN
      return(doseX.tj)
    }
    long2.c.spl<-split(long2.c, long2.c[,Select.Var])
    dose1.clean<-lapply(long2.c.spl, format.ds, var1='pcv1.aged', doseN=1)
    dose2.clean<-lapply(long2.c.spl, format.ds, var1='pcv2.aged', doseN=2)
    dose3.clean<-lapply(long2.c.spl, format.ds, var1='pcv.boost.aged', doseN=3)
    dose1.clean<-do.call('rbind.data.frame', dose1.clean)
    dose2.clean<-do.call('rbind.data.frame', dose2.clean)
    dose3.clean<-do.call('rbind.data.frame', dose3.clean)
    
    dose.comp<-rbind.data.frame(dose1.clean,dose2.clean,dose3.clean)
    dose.comp$agem<-dose.comp$aged/30.3
    dose.comp$agem[dose.comp$agem<0]<-0
    
    dose.comp<-dose.comp[order(dose.comp[,Select.Var],dose.comp$doseN),]
    
    dose.comp[,Select.Var]<-factor(dose.comp[,Select.Var])
    Stratas<-unique(dose.comp[,Select.Var])
    
    cols<-c('#e41a1c','#377eb8','#4daf4a','#984ea3','#ff7f00')
    form2<-as.formula(paste0('Surv(dose.comp$agem,dose.comp$doseX)~doseN+',Select.Var))
    dosemod<- survfit(form2, data=dose.comp)
    
    surv.summary<-surv_summary(dosemod,data=dose.comp)
    surv.summary.spl<-split(surv.summary, paste(surv.summary$doseN, surv.summary[,Select.Var]))
    uptake<-1-sapply(surv.summary.spl, function(x) x$surv[length(x$surv)])
    uptake
    
    form3<-as.formula(paste0('time~doseN~',Select.Var))
    surv.summary.c<-1-acast(surv.summary[,c('time','doseN',Select.Var,'surv') ],form3 )
    time.m<-as.numeric(dimnames(surv.summary.c)[[1]])
    target.time<-c(1.5,3.5, 9) #when should the vaccine be given?
    par(mfrow=c(1,3), mar=c(4,4,1,1))
    for(i in 1:3){
      matplot(time.m,surv.summary.c[,i,], type='p', cex=0.25, pch=16,col=cols,bty='l', xlab='Age, months', ylab='Proportion vaccinated',lty=1, ylim=c(0,1))
      abline(v=target.time[i], col='gray', lty=2)
    }
  p1.base<- recordPlot()
 out.obj<-list('plots'=p1.base,'uptake'=uptake,'Strata'=Var.Levels, 'summary.table'=strata.tab1)
}

```

## Variability in uptake by Block
```{r,fig.width=7, fig.height=2.5}
block.analysis<-stratified.function('Block')
block.analysis$plots
block.analysis$summary.table
block.analysis$uptake
```

## Variability in uptake by sex
```{r,fig.width=7, fig.height=2.5}
sex.analysis<-stratified.function('Sex')
sex.analysis$plots
sex.analysis$summary.table
sex.analysis$uptake
```




## 3 dose coverage for 12-24 month old kids
```{r}

ds.cov<-cbind.data.frame('pcv3'=pcv3,'pcv2'=pcv2,'pcv1'=pcv1, 'boost.date'=as.Date(ds1$PCV.Booster.Date,'%Y-%m-%d'),
                         'dose1.date'=as.Date(ds1$PCV1.Date,'%Y-%m-%d'),
                         'dose2.date'=as.Date(ds1$PCV2.Date,'%Y-%m-%d'),
                         'id'=1:length(pcv3), 'dob'=as.Date(ds1$DOB) )
monthdateseq<-seq.Date(from=as.Date('2018-03-01'),to=as.Date('2019-10-01'), by='month')
id.seq<-1:nrow(ds.cov)
id.ds<-expand.grid(monthdateseq,id.seq)
names(id.ds)<- c('monthdate', 'id')
ds.cov2<-merge(ds.cov, id.ds, by='id', all=T)

ds.cov2$dose1.vax.status.date<-0
ds.cov2$dose1.vax.status.date[ds.cov2$dose1.date<=ds.cov2$monthdate] <-1
ds.cov2$dose2.vax.status.date<-0
ds.cov2$dose2.vax.status.date[ds.cov2$dose2.date<=ds.cov2$monthdate] <-1
ds.cov2$dose3.vax.status.date<-0
ds.cov2$dose3.vax.status.date[ds.cov2$boost.date<=ds.cov2$monthdate] <-1
ds.cov2$agem<- as.numeric((ds.cov2$monthdate - ds.cov2$dob))/30.3
ds.cov2.12.24m<-ds.cov2[ds.cov2$agem>=12 & ds.cov2$agem<24,]
cov.date<-aggregate(ds.cov2.12.24m[,c('dose1.vax.status.date','dose2.vax.status.date','dose3.vax.status.date')], by=list('monthdate'=ds.cov2.12.24m$monthdate), FUN=mean)
trans.col<- rgb(0,0,0, alpha=0.1)
plot(cov.date$monthdate, cov.date$dose1.vax.status.date, type='l', bty='l', xaxt='n', xlab='Date', ylab='Uptake (12-24m)', ylim=c(0,1), las=1)
abline(h=c(0.2, 0.4, 0.6,0.8,1.0), col=trans.col, lty=1)
points(cov.date$monthdate,cov.date$dose1.vax.status.date, type='l' , lty=1, lwd=2)
points(cov.date$monthdate,cov.date$dose2.vax.status.date, type='l' , lty=2, lwd=2)
points(cov.date$monthdate,cov.date$dose3.vax.status.date, type='l' , lty=3, lwd=2)
axis(side=1, at=as.Date(c('2018-03-01','2018-09-01', '2019-03-01', '2019-09-01')), labels=c('Mar 2018','Sep 2018', 'Mar 2019', 'Sep 2019'))

```


Age of kids who get 1st dose
```{r}
ds1$age.1st.dose<-  as.numeric(as.Date(ds1$PCV1.Date, "%Y-%m-%d") - as.Date(ds1$DOB, "%Y-%m-%d"))
ds1$age.2nd.dose<-  as.numeric(as.Date(ds1$PCV2.Date, "%Y-%m-%d") - as.Date(ds1$DOB, "%Y-%m-%d"))
ds1$age.3rd.dose<-  as.numeric(as.Date(ds1$PCV.Booster.Date, "%Y-%m-%d") - as.Date(ds1$DOB, "%Y-%m-%d"))

hist(ds1$age.1st.dose/30.3)
hist(ds1$age.2nd.dose/30.3)
hist(ds1$age.3rd.dose/30.3)

hist(ds1$age.1st.dose[year(as.Date(ds1$PCV1.Date, "%Y-%m-%d"))==2018 ] /30.3, main='Dose 1 in 2018')

par(mfrow=c(1,3), mar=c(4,2,1,1))
hist(ds1$age.1st.dose[year(as.Date(ds1$PCV1.Date, "%Y-%m-%d"))==2019 ] /30.3, main='Dose1, 6w', xlim=c(0,14), breaks=30, freq=F, xlab='Age(m)')
abline(v=1.5, lty=2, col='red')
hist(ds1$age.2nd.dose[year(as.Date(ds1$PCV2.Date, "%Y-%m-%d"))==2019 ]/30.3, xlim=c(0,14),main='Dose 2, 14w', breaks=30, freq=F, xlab='Age(m)')
abline(v=3.5, lty=2, col='red')
hist(ds1$age.3rd.dose[year(as.Date(ds1$PCV.Booster.Date, "%Y-%m-%d"))==2019 ]/30.3, xlim=c(0,14),main='Dose 3, 9m', breaks=30, freq=F, xlab='Age(m)')
abline(v=9, lty=2, col='red')

```

#Vaccine timeliness among KB population
As it is currently, this shows there are a bunch of kids getting roughly schedule, and a bunch who are getting catch up. Need to strtaify based on whether kids were eligible for the routine doses or not. If Routine immunization started in april 2018, then kids born after, say Feb 1, 2018 would be eligible for routine schedule. As analysis above shows, it takes a few months before really see high dose1 uptake. kids born after july or aug have good uptake of first dose

```{r timeliness, echo=FALSE}
age.dose1<-  as.Date(ds1$PCV1.Date[which(pcv1==1 &routine.eligible.flag==1)])- as.Date(ds1$DOB[which(pcv1==1 & routine.eligible.flag==1)])

#plot(age.dose1, as.Date(ds1$DOB[which(pcv1==1 & routine.eligible.flag==1)]), bty='l', main='Later Age of 1st dose, earlier DOB', sub='suggests catch up doses')

age.dose2<-  as.Date(ds1$PCV2.Date[which(pcv2==1 &routine.eligible.flag==1)])- as.Date(ds1$DOB[which(pcv2==1 &routine.eligible.flag==1)])

age.dose3<-  as.Date(ds1$PCV.Booster.Date[which(pcv3==1 &routine.eligible.flag==1)])- as.Date(ds1$DOB[which(pcv3==1 &routine.eligible.flag==1)])

print('age in days of eligible kids who received PCV 1st or second dose')
par(mfrow=c(1,2))
hist(as.numeric(age.dose1)/30.2, main=median(as.numeric(age.dose1))/30.3)
hist(as.numeric(age.dose2)/30.3, main=median(as.numeric(age.dose2))/30.3)
#hist(as.numeric(age.dose3), main=median(as.numeric(age.dose3))/30)
par(mfrow=c(1,1))



#Time between first and second dose
#THESE VECTORS HAVE DIFFERENT LENGTH
#dose.gap1.2<-age.dose2-age.dose1#
#hist(as.numeric(dose.gap1.2), main=paste0('Median gap ',median(dose.gap1.2), ' days'))

```

#Next look at reports of respiratory illness among the KB cohort
The vast majority have missing information. Is this because the question wasn't asked?
```{r resp_illness, echo=FALSE}
pneu1<-ds1$
pneu1[is.na(pneu1)]<-999
table(pneu1)

```

#And how about in the PCTS system?
In what age group are 'pneumonia cases' calculated? Is tis kids only? Including adults?
```{r pcts, echo=FALSE}
pc1<-read.csv('../kb_data/Updated_PCTS_Count_Data.csv')
  # "Infant.death.Pneumonia..1.11.Months" ;"Infant.death.Pneumonia..1.5..Yrs.";"Pneumonia.Cases"  ;"Month"    
pc1$Block <- autoFill(pc1$Block)
table(pc1$Block)

overall<-pc1[pc1$S.NO.=='Total',]
overall$month2<-paste0('01-',overall$Month)
overall$date<-as.Date(overall$month2, format = '%d-%b%Y')
par(mfrow=c(1,1))
plot(overall$date, overall$Pneumonia.Cases, bty='l', type='l', main='Pneumonia cases')
plot(overall$date, overall$Infant.death.Pneumonia..1.11.Months, bty='l', type='l', main='Pneumonia Death 1-11m')
plot(overall$date, overall$Infant.death.Pneumonia..1.5..Yrs., bty='l', type='l', main='Pneumonia deaths 1-5y')
par(mfrow=c(1,1))
  
```

Look at PCTS data just for blocks present in Khushi baby data
```{r}
pc2<-pc1[pc1$Block %in% blocknames,]
pc2$month2<-paste0('01-',pc2$Month)
pc2$date<-as.Date(pc2$month2, format = '%d-%b%Y')
#Aggregate by block
pc3<-aggregate(pc2[,c("Pneumonia.Cases","Infant.death.Pneumonia..1.11.Months","Infant.death.Pneumonia..1.5..Yrs.")], by=list('block'=pc2$Block, 'date'=pc2$date), FUN=sum)
pc3$block<-factor(pc3$block)

plot(pc3$date, pc3$Pneumonia.Cases, col=cols[pc3$block])
plot(pc3$date, pc3$Infant.death.Pneumonia..1.11.Months, col=cols[pc3$block])
plot(pc3$date, pc3$Infant.death.Pneumonia..1.5..Yrs., col=cols[pc3$block])


```

Map. Need to overlay a shapefile
```{r}
ds1$Updated.Lat[ds1$Updated.Lat<20]<-NA
ds1$Updated.Long[ds1$Updated.Long<60]<-NA

ds1$vax.12.23.boost<-NA
ds1$vax.12.23.boost[(as.Date('2019-10-21')-as.Date(ds1$DOB))>365 &   as.Date(ds1$DOB)>=as.Date('2018-03-01') ]<-0
ds1$vax.12.23.boost[ds1$PCV.Booster.Given.Denied=='Given'] <-1 
cols2<-c(rgb(0,0,0, alpha=0.1),rgb(1,0,0, alpha=0.1))
plot(ds1$Updated.Long, ds1$Updated.Lat, col=cols2[(ds1$vax.12.23.boost+1)], pch=16)

```


